A First Effort for a Distributed Segment-based Approach on Self-Assembled Nano Networks

### 2014-02-12 PAPER/disr.tex
General text reorganization/fixes

%------------------------------------------------------------------------------
\section{Preliminary concepts}
%------------------------------------------------------------------------------
\label{sec:disr_concepts}

\subsection{Basic Idea}
The main concept behind of the original SR approach is the
\emph{Segment}: a Segment is basically a path of consecutive node
and links. All segments start with a link attached to a node which
belongs to different segment, ending with a link also attached to a
node belonging to a different segment. A exception is the first
segment estabilished in the network, called \emph{starting segment} which is  a loop
beginning/ending on a particular node defined as bootstrap node.

The idea is to partition the network in a set of disjoint segments, and then
placing a turn restriction within each segment. It has been proved~\cite{mejia_ipdps06}
that such a set of turn prohibitions guarantees dealock freedom and
while preserving connectivity of the network.

Describing the execution phases of DiSR from a top level point-of-view
is basically similar to the classic SR approach. This can be useful to
give an initial idea of the approach, however, some essential
differences should be pointed out. First of all no centralized entity is
globally aware of what is going on: the status of the DiSR execution
is collectively distributed among the nodes; second, no defect map and/or
topology graph is available as input, thus, the topology has to be
discovered \emph{while} segments are created; finally, at the end of the
execution, no segment list is created: Each node is only aware of
belonging to some segment, ignoring the presence of other nodes
in the same segment and even the presence of other segments in the
network.

Roughly speaking, the execution of can be described as the execution
of the following phases:
\begin{itemize}
\item Injection of the DiSR process from upper layer to set a bootstrap
node
\item Bootstrap node broadcasting to create the first segment of the subnet
\item Parallel requests starting from assigned node to discover other segments
%\item Turn proibition in each confirmed segment
%\item Draining of remaining packet due timeouts
\end{itemize}

Once again, it should be pointed out that nowhere in the network there
is such a global vision of these exection phases. A more detailed
description of the exection model at each node is discussed in the
next section.

\subsection{Packets types Required}

The DiSR approach works with a distributed mechanism which is build
upon an exchange of small packets containing the following fields:
\begin{itemize}
\item{packet\_type}: encodes the meaning of the control message respect
to the whole DiSR process.
\item{seg\_ID}: the id of the segment associated to the DiSR control message
\item{src\_id}: the id of the node that initially originated the
packet
\end{itemize}

As regards packet\_type, we can have the following control packets:
\begin{itemize}
\item{\texttt{STARTING\_SEGMENT\_REQUEST}}: Injected by the bootstrap
node to when searching for the first segment. 
\item{\texttt{STARTING\_SEGMENT\_CONFIRM}}: used when estabilishing
the starting  segment. 
\item{\texttt{SEGMENT\_REQUEST}}: used to search candidates for a segment (not the
first)
\item{\texttt{SEGMENT\_CONFIRM}}: used to estabilish a segment. 
\item{\texttt{SEGMENT\_CANCEL}}: used to cancel the process of searching a segment along a
specific link.
\end{itemize}

A quantitative analysis of the resources needed to implement these
fields is presented in Section~\ref{sec:implementation} when
discussing the impact of the DiSR control logic and storage on
hardware implementation of the node.

\subsection{DiSR Data Structures}
\label{ssec:disr_dstruct}

@@

The \emph{LED} is like a snapshot of the DiSR algorithm at each node,
consisting in the following variables:

\begin{itemize}
\item{\emph{segID}}:a value used to specify the segment to which the
node has been assigned or is candidate for being assigned.
\item{\emph{visited}}: a boolean value. When \emph{true}
a \emph{segID} different from \texttt{NULL} specifies the segment 
to which the node has been assigned.

% TODO: the following is true, but complicates things at this stage
%Note that a value different from \emph{NULL} does not
%necessarily mean the node has been assigned to some segment (e.g. when
%searching starting segment is set \emph{visited} by default even if not yet
%assigned to any segment).

\item{\emph{tvisited}}: a boolean value. If \emph{true}, the node is
being considered as candidate for a segment, and the \emph{segID} value
specifies the segment id for which the node is candidate. 
\item{\emph{link\_visited[]}}: an array
of values representing information about attached links, that is, the
segID of the segment owning each link. When \texttt{NULL}, the corresponding link has not yet been
assigned.
%TDB: do we need to introduce some value for the
%“bridge” status ? (e.g. a node connecting the terminal node of a
%subnet with the starting one of another) 
\item{\emph{link\_tvisited[]}}: an array of
values representing information about attached links, that is, the \emph{segID} of
the segment for which the link is candidate. When \texttt{NULL}, the link is
currently not \emph{tvisited}.  
\end{itemize}
%\item{\emph{starting}}: boolean, \emph{true} if the node is a
%starting node from which the whole process was initiated.
%\end{itemize}

@@

\begin{itemize}
\item{\emph{Free}}: a node that has  not been yet considered  by the DiSR algorithm.
The node is not marked as \emph{visited}/\emph{tvisited}.
\item{\emph{Bootstrap}}: a node which has been explicitly set as bootstrap node from
an upper layer via. 
% TODO: this and all the other details on the items below should be
% checked and moved elsewhere for a future long paper

%This is required because a
%\texttt{STARTING\_SEGMENT\_REQUEST} packet should be injected in order to start
%the whole process.
\item{\emph{ActiveSearching}}: a node from which a find process of new segments has
been started and not yet cancelled or confirmed. 
%This happens when all
%the following conditions are matched: the node is marked as \emph{visited},
%that is \emph{visited} = TRUE 
%TODO: \emph{segID} is set to some value reflecting
%the segment request sent there is a link that is being investigated by
%the node, that is a tvisited\_link marked with the same id of the node 
\item{\emph{Candidate}}: a node currently candidate for belonging to some segment
with id segid, not being itself the node from which the find process
was started.

%The following conditions should be matched: \emph{tvisited} =
%true, with segID=X different from node's id.  Two links set as
%\emph{link\_tvisited} with id X. Let’s refer them as  links 'i' and 'j'. Then
%link\_tvisited[i] was set = X when the find process from an adjacent
%node reached the current node, while link\_tvisited[j] has been set to
%X when the node itself started to investigate its free links for the
%segment request with id X. Note that while investigating its attached
%links, if the find process fails along path 'j', the current node can
%investigate another free node 'k' if suitable.  
\item{\emph{CandidateStarting}}: same as above, but the node is currently being considered as candidate
for starting segment.

%Two main differences: Can confirm
%\texttt{STARTING\_SEGMENT\_CONFIRM} with the same \emph{segID} when receiving a normal
%\texttt{SEGMENT\_REQUEST}: Overwrites its status setting it to \emph{Candidate} for the
%segID frees the links previously set as \emph{tvisited} during flooding
\item{\emph{Assigned}}: a node for which the segment has been determined.  The
segment \emph{segID} attribute value is set to some id X different from \texttt{NULL}.
%The \emph{visited} value is \emph{true} NOTE: an \emph{Assigned} state is a quick temporary
%state, since it becomes \emph{ActiveSearching} if a not (\emph{visited}/tvisited)
%link is suitable for searching new segments (see next\_visited
%procedure on paper) 
\end{itemize}

\subsection{Sketch of DiSR}

\begin{itemize}

\item{\textbf{Injecting bootstrap request}}: all nodes have a initial
\emph{DBS} status \emph{Free}, except for a
 node with status \emph{Bootstrap}, set by some signal from an
upper layer via. When starting, this bootstrap node sets itself as \emph{visited} and
changes its status to \emph{ActiveSearching}, injecting a
\texttt{STARTING\_SEGMENT\_REQUEST} across one of its free links.

\item{\textbf{Flooding}}: Each node receiving a \texttt{STARTING\_SEGMENT\_REQUEST},
if not yet (\emph{visited}/\emph{tvisited}),  forwards it to its free links using a flooding mechanism.
Each of these links is then marked as \emph{tvisited} with the segment id
associated to the request. Note that a node that has already received a
\texttt{STARTING\_SEGMENT\_REQUEST} packet can simply ignore further packets
associated to the same request, having already contributed to the
flooding. 
%Note also that each of these packets has a max\_segment\_hops
%field to prevent packets from travelling undefinitively.

\item{\textbf{Confirming the Starting Segment}}: when this \texttt{STARTING\_SEGMENT\_REQUEST} packet reaches
the starting node (from a different link, of course), the starting
segment is found. Then the starting node sends back a \texttt{STARTING\_SEGMENT\_CONFIRM}
packet with the same id along the link from which it received the
\texttt{STARTING\_SEGMENT\_REQUEST}. Each node do same by changing its own status
to \emph{Assigned}. So the confirmation packet is sent back from
node to node and the starting segment is created.

Note: A node knows to be \emph{Candidate} due a
\texttt{STARTING\_SEGMENT\_REQUEST}.  Then, when receiving a simple
(not starting) \texttt{SEGMENT\_REQUEST} , it can simply cancel its
previous \emph{Candidate} status, since this second
request means that the starting segment has already been found.

@@

\item{\textbf{Injecting other request}}: Each node in the \emph{Assigned} status can initiate a search for a
segment, by sending a \texttt{SEGMENT\_REQUEST} across one of its free
links. Note that, since this is not the starting segment, in this case
the packet should not return back to the initiator, but just another visited
node
\item{\textbf{Setup of a segment}}: The find process is successfull
when an \emph{Assigned} node receives the \texttt{SEGMENT\_REQUEST} packet. Then, a \texttt{SEGMENT\_CONFIRM} is
sent back along the path that originated the request.  
\item{\textbf{Failing while searching a segment}}: a node received a \texttt{SEGMENT\_REQUEST} packet but
matched one of the following conditions: the node is free but has no
more suitable free links (can’t forward the \texttt{SEGMENT\_REQUEST}) the node
is \emph{tvisited} candidate for another find process the request packet
exceeded the max\_segment\_hops limit.  In all these cases the node
sends back a \texttt{SEGMENT\_CANCEL} along the proper link modifying its state
to free. Each of the subsequent nodes will forward this \texttt{SEGMENT\_CANCEL}
only if no more free links can be explored starting from them.
\end{itemize}

%\subsection{Intra-node vs Inter-node parallelism}
%In the processes described above, we assumed that each node in the
%READY\_SEARCHING state can start a find segment process by injecting a
%\texttt{SEGMENT\_REQUEST}. However, a more accurate decision when defining DiSR
%should be made in order to decide whether or not the node must
%actually perform this action. 
%The point above is is strictly related to the general question: which
%level of “parallelism” should we allow in DiSR ? The adopted approach
%is that, although the nature of DiSR itself is intrinsicly parallel,
%the use of parallelism makes things work in a more complex way. In
%other words, DiSR is parallel when needed, but it does not exploit
%parallelism as an “improving feature”. When not needed, things should
%be serialized. For example: when a node is “ready searching” could
%start several find segment process associated to the same segment, one
%for each free link. But serializing this by investigating the free
%links in order could be a simplest solution. We refer to this saying
%that we avoid intra-node parallelism.  Note that while intra-node
%parallelism can be avoided, trying to avoid inter-node parallelism
%could be more complex than allowing the parallelism itself. Imagine
%for example the effort of trying to coordinate nodes so that an unique
%finding node process is actually running in the whole subnet. Thus, in
%contrast to the intra-node parallelism, the inter-node parallelism is
%a structural property of the DiSR algorithm and should not be avoided.
%------------------------------------------------------------------------------

@@

\begin{figure}
  \centering
    \includegraphics[width=0.30\textwidth]{pictures/state_machine.eps}
  \caption{DiSR node execution model}
  \label{fig:dna_tag}
\end{figure}

@@

%TODO REPEAT!!
%When a node discovers that the path that includes
%the node itself is no more suitable, a \texttt{SEGMENT\_CANCEL} is sent back
%across the link from which the \texttt{SEGMENT\_REQUEST} was received.  Possible
%reasons for that: The node it’s already candidate for another segment
%The node it’s free but can’t find a free link (according the
%cyclelinks timeout) The node it’s \emph{tvisited} for the same request id,
%previously forwarded that request along some free direction, but now
%it’s receiving a a cancel request and has no more free links

>>>>

%------------------------------------------------------------------------------
\section{\disr{} Sketch}
%------------------------------------------------------------------------------

\label{sec:disr_concepts}
The main concept behind the kind of turn prohibition we want to
achieve is the \emph{Segment}. A segment is basically a path of consecutive
nodes and links, starts with a link attached to a node of a different
segment and ending with a link attached to a
node belonging another. In other words, a segment
a path between connecting two other different segments. The example
in Figure~\ref{fig:segments} shows the segment with id $1.4$, starting
with a link connected to the node $9$, the going trought nodes
$4$,$3$,$2$ and ending with the link connected to the node $1$. Other
segments depicted are $9.2$ (link,node $10$,link) and $9.3$
(link from $9$ ,$8$,link,$7$,link,$6$,link,$5$,link to the $1$). Note
also that start/ending link can coincide and a segment can also be
formed by one single link (e.g. in $7.2$ or $5.4$).
An exception to the rule that a segment must connect two other
segments is the first segment established in the network, called
\emph{starting segment} which is a loop beginning/ending on the same
node.

\begin{figure}
\centering
    \includegraphics[width=0.25\textwidth]{pictures/network_slice.eps}
  \caption{Example of Segments}
  \label{fig:segments}
\end{figure}

A detailed description of the \disr{} execution model as depicted in
the Figure~\ref{fig:disr_events} will be discussed in
Subsection~\ref{sub:phases}. Roughly speaking, the execution of can be described as the execution
of the following phases: 
\begin{enumerate}
\item Injection of the \disr{} process from an upper layer to set a \emph{bootstrap
node} 
\item Broadcasting and confirmation to create the first segment of the subnet
\item Parallel requests starting from assigned node to discover other
segments. 
\item Turn proibition in each confirmed segment.
\item Draining of remaining packet due timeouts
\end{enumerate}

\section{Detailed Description of \disr{}}
\subsection{Message Types Required}

The \disr{} approach works with a distributed mechanism which is build
upon an exchange of small packets containing the following fields:
\begin{itemize}
\item \emph{packet\_type}: encodes the meaning of the control message
\item \emph{seg\_ID}: the id of the segment associated to the \disr{} control message
\item \emph{src\_id}: the id of the node that originated the packet
\end{itemize}

As regards \emph{packet\_type} field, we can have the following control packets:
\begin{itemize}
\item{\texttt{STARTING\_SEGMENT\_REQUEST}}: Injected by the bootstrap
node to when searching for the first segment. 
\item{\texttt{STARTING\_SEGMENT\_CONFIRM}}: used when establishing
the starting  segment. 
\item{\texttt{SEGMENT\_REQUEST}}: used to search candidates for a segment (not the
first)
\item{\texttt{SEGMENT\_CONFIRM}}: used to establish a segment. 
\item{\texttt{SEGMENT\_CANCEL}}: used to cancel the process of searching a segment along a
specific link.
\end{itemize}

\subsection{\disr{} Data Structures}
\label{ssec:disr_dstruct}

@@

The \emph{LED} is like a snapshot of the \disr{} algorithm at each node,
consisting in the following variables:
\begin{itemize}
\item{\emph{segID}}:a value used to specify the segment to which the
node has been assigned or is candidate for being assigned.
\item{\emph{visited}}: a boolean value. When \emph{true}
a \emph{segID} different from \texttt{NULL} specifies the segment 
to which the node has been assigned. 
% TODO: the following is true, but complicates things at this stage
%Note that a value different from \emph{NULL} does not
%necessarily mean the node has been assigned to some segment (e.g. when
%searching starting segment is set \emph{visited} by default even if not yet
%assigned to any segment). 
\item{\emph{tvisited}}: a boolean value. If \emph{true}, the node is
being considered as candidate for a segment, and the \emph{segID} value
specifies the segment id for which the node is candidate. 
\item{\emph{link\_visited[]}}: an array
of values representing information about attached links, that is, the
segID of the segment owning each link. When \texttt{NULL}, the corresponding link has not yet been
assigned.
%TDB: do we need to introduce some value for the
%“bridge” status ? (e.g. a node connecting the terminal node of a
%subnet with the starting one of another) 
\item{\emph{link\_tvisited[]}}: an array of
values representing information about attached links, that is, the \emph{segID} of
the segment for which the link is candidate. When \texttt{NULL}, the link is
currently not \emph{tvisited}.  
\end{itemize}
%\item{\emph{starting}}: boolean, \emph{true} if the node is a
%starting node from which the whole process was initiated.
%\end{itemize}

@@

\begin{itemize}
\item{\emph{Free}}: a node that has  not been yet considered  by the \disr{} algorithm.
\item{\emph{Bootstrap}}: a node which has been explicitly set as bootstrap node from
an upper layer via. 
\item{\emph{ActiveSearching}}: a node from which a find process of new segments has
been started and not yet cancelled or confirmed. 
%This happens when all
%the following conditions are matched: the node is marked as \emph{visited},
%that is \emph{visited} = TRUE 
%TODO: \emph{segID} is set to some value reflecting
%the segment request sent there is a link that is being investigated by
%the node, that is a tvisited\_link marked with the same id of the node 
\item{\emph{Candidate}}: a node currently candidate for belonging to some segment
, but not being itself the node from which the find process was started.

%The following conditions should be matched: \emph{tvisited} =
%true, with segID=X different from node's id.  Two links set as
%\emph{link\_tvisited} with id X. Let’s refer them as  links 'i' and 'j'. Then
%link\_tvisited[i] was set = X when the find process from an adjacent
%node reached the current node, while link\_tvisited[j] has been set to
%X when the node itself started to investigate its free links for the
%segment request with id X. Note that while investigating its attached
%links, if the find process fails along path 'j', the current node can
%investigate another free node 'k' if suitable.  
\item{\emph{CandidateStarting}}: same as above, but the node is currently being considered as candidate
for starting segment. 
%Two main differences: Can confirm
%\texttt{STARTING\_SEGMENT\_CONFIRM} with the same \emph{segID} when receiving a normal
%\texttt{SEGMENT\_REQUEST}: Overwrites its status setting it to \emph{Candidate} for the
%segID frees the links previously set as \emph{tvisited} during flooding
\item{\emph{Assigned}}: a node for which the segment has been determined.  
%The \emph{visited} value is \emph{true} NOTE: an \emph{Assigned} state is a quick temporary
%state, since it becomes \emph{ActiveSearching} if a not (\emph{visited}/tvisited)
%link is suitable for searching new segments (see next\_visited
%procedure on paper) 
\end{itemize}

A quantitative analysis of the resources needed to implement these
fields is presented in Section~\ref{sec:implementation} when
discussing the impact of the \disr{} control logic and storage on
hardware implementation of the node.

\subsection{Execution Model}
\label{sub:phases}

The \disr{} control packets described in the previous subsection trigger
node events, which eventually change $DBS$ status according to the
current value of $LED$ variables and the type of the packet received.
The main \disr{} phases, together with a reference to the particular
status transition involved are described in the following. As a
convention, we used single letters to refer to one of the \emph{DBS} transitions in
Figure~\ref{fig:state_machine} while the phases are visually
represented in~\ref{fig:disr_events}.

\begin{figure}
  \centering
    \includegraphics[width=0.25\textwidth]{pictures/dbs_updated.eps}
  \caption{\disr{} node execution model}
  \label{fig:state_machine}
\end{figure}
\begin{itemize}

\begin{figure*}
\centering
\begin{tabular}{cccc}
\includegraphics[width=0.23\textwidth]{pictures/seq01.eps} & 
\includegraphics[width=0.23\textwidth]{pictures/seq02.eps} &
\includegraphics[width=0.23\textwidth]{pictures/seq03.eps} & 
\includegraphics[width=0.23\textwidth]{pictures/seq04.eps} \\
(a) & (b) & (c) & (d) \\
\includegraphics[width=0.23\textwidth]{pictures/seq05.eps} &
\includegraphics[width=0.23\textwidth]{pictures/seq06.eps} &
\includegraphics[width=0.23\textwidth]{pictures/seq07.eps} & 
\includegraphics[width=0.23\textwidth]{pictures/seq08.eps} \\
(e) & (f) & (g) & (h) \\
\end{tabular}
\caption{\disr{} (a) and RPF (b) node coverage}
\label{fig:disr_events}
\end{figure*}

\item{\textbf{Injecting bootstrap request}}: all nodes have an initial
\emph{DBS} status set to \emph{Free}, except for a
 node with status \emph{Bootstrap}, set by some control signal from an
upper layer via $(a)$. When starting, this bootstrap node 
changes its status to \emph{CandidateStarting} $(b)$, injecting a
\texttt{STARTING\_SEGMENT\_REQUEST} across one of its free links
(Figure~\ref{fig:disr_events}-a).

\item{\textbf{Bootnode Broadcasting}}: Each node receiving a \texttt{STARTING\_SEGMENT\_REQUEST},
when $Free$, forwards it to its free links using a flooding mechanism and
becoming $CandidateStarting$ $(g)$.
Each of its free links is then marked as \emph{tvisited} with the segment id
associated to the request. Note that a node that has already received a
\texttt{STARTING\_SEGMENT\_REQUEST} packet can simply ignore further packets
associated to the same request, having already contributed to the
flooding (Figure~\ref{fig:disr_events}-b).
%Note also that each of these packets has a max\_segment\_hops
%field to prevent packets from travelling undefinitively.

\item{\textbf{Confirming the Starting Segment}}: when a \texttt{STARTING\_SEGMENT\_REQUEST} packet reaches
the bootstrap node (from a different link, of course), the starting
segment is found. Then the bootstrap node sends back a \texttt{STARTING\_SEGMENT\_CONFIRM}
packet along the link from which it received the
\texttt{STARTING\_SEGMENT\_REQUEST} and becomes $Assigned$ $(l)$. Each
node receiving the confirm do same by changing its own status
to \emph{Assigned}(Figure~\ref{fig:disr_events}-c). So the confirmation packet is sent back from
node to node and the starting segment is created (Figure~\ref{fig:disr_events}-d).

@@

\item{\textbf{Injecting Segment Requests}: Each node in the
\emph{Assigned} status can initiate a search for a segment, by sending
a \texttt{SEGMENT\_REQUEST} across one of its free links $(c)$ (see
Figure~\ref{fig:disr_events}-e). 
Note that each \emph{Assigned} note can inject the request just after
becoming assigned, i.e. it is not needed that the whole confirmation
process in completed (Figure~\ref{disr_events}-f).
Note also that nodes previosly set as \emph{CandidateStarting}, when receiving 
a \texttt{SEGMENT\_REQUEST} can simply cancel their
status and set as $Candidate$ for that request, since this new
request means that the first starting segment has already been found $(h)$. 
\item{\textbf{Segment Confirmation}}: The find process is successful
when an already \emph{Assigned} node receives the
\texttt{SEGMENT\_REQUEST} packet. Then, a \texttt{SEGMENT\_CONFIRM} is
sent back along the path that originated the request
(Figure~\ref{fig:disr_events}-g), while the
node remain $Assigned$ (since it could confirm more segments). Each node
previously set as $Candidate$ for that segment id, when receiving the
a confirmation packet changes its status to $Assigned$ $(m)$ and forwards back
the same confirmation until it reaches the initiator of the request,
which then changes from $ActiveSearching$ to $Assigned$ $(d)$.

\item{\textbf{Failing while searching a segment}}: this happen when a node received a \texttt{SEGMENT\_REQUEST} packet but
matches one of two the following conditions: first, the node is $Free$ but has no
more suitable free links (thus can’t forward the
\texttt{SEGMENT\_REQUEST})$(f)$; the node
is already \emph{Candidate} with another segment id, i.e. another find
process.  In all these cases the node
sends back a \texttt{SEGMENT\_CANCEL} along the incoming link (see at
the bottom of Figure~\ref{fig:disr_events}).
%Each of the subsequent nodes will forward this \texttt{SEGMENT\_CANCEL}
%only if no more free links can be explored starting from them.
\end{itemize}

% TODO: FIX and RE-enable in long paper

@@

@@

%TODO REPEAT!!
%When a node discovers that the path that includes
%the node itself is no more suitable, a \texttt{SEGMENT\_CANCEL} is sent back
%across the link from which the \texttt{SEGMENT\_REQUEST} was received.  Possible
%reasons for that: The node it’s already candidate for another segment
%The node it’s free but can’t find a free link (according the
%cyclelinks timeout) The node it’s \emph{tvisited} for the same request id,
%previously forwarded that request along some free direction, but now
%it’s receiving a a cancel request and has no more free links
\subsection{Intra-node vs Inter-node parallelism}
In the processes described above, we assumed that each node in the
\emph{Assigned} status can start a find segment process by injecting a
\texttt{SEGMENT\_REQUEST}. However, a more accurate decision when
defining \disr{} should be made in order to decide whether or not the
node must actually perform this action.  The point above is is
strictly related to the general question: which level of “parallelism”
should we allow in \disr{} ? The adopted approach is that, although
the nature of \disr{} itself is intrinsicly parallel, the use of
parallelism makes things work in a more complex way. In other words,
\disr{} is parallel when needed, but it does not exploit parallelism
as an “improving feature”. When not needed, things should be
serialized. For example: when a node is “ready searching” could start
several find segment process associated to the same segment, one for
each free link. But serializing this by investigating the free links
in order could be a simplest solution. We refer to this saying that we
avoid intra-node parallelism.  Note that while intra-node parallelism
can be avoided, trying to avoid inter-node parallelism could be more
complex than allowing the parallelism itself. Imagine for example the
effort of trying to coordinate nodes so that an unique finding node
process is actually running in the whole subnet. Thus, in contrast to
the intra-node parallelism, the inter-node parallelism is a structural
property of the \disr{} algorithm and should not be avoided.
%------------------------------------------------------------------------------

### 2014-02-12 PAPER/introduction.tex
General text reorganization/fixes

\section{Related Works}
\label{sec:related_works}
Several works~\cite{sancho2002, skeie2002, skeie2004, koibuchi2003} have been presented addressing the problem of topology
agnostic routing. While they show interesting
performances, the requirement for virtual channels resources 
to avoid deadlock makes them not suitable for the specific DNA self
assembled networks we are assuming as background.
Other topology independent algorithms~\cite{schroeder1991, koibuchi2001, cherkasova1996} do not use virtual channels to
achieve deadlock freedom, but turn prohibitions. 
In particular, authors of~\cite{Patwardhan05evaluatingthe} exploits the creation of a spanning tree of the
topology, then placing bidirectional restrictions by avoiding a packet
to traverse the same link in both up and down directions.
While the hierarchical nature of this approach can lead to uneven traffic
distribution, with many packets traversing upper links (near to the
root), this is quite acceptable in classical wide area networks
topologies with a limited number of nodes. Other approaches such as
FX~\cite{sancho2000} mitigate this
issue, but the set of turn restrictions is still prefixed,
strictly depending on the particular tree root selected. 
Other solutions try to approach the issue of irregular
topologies by limiting the number~\cite{dally1994, duato1997, gomez2004, koibuchi2008} or the
location~\cite{zhang2008, sui2000, flich2008, liu2011} of missing links. This
restriction is clearly unacceptable given the hight-defect rates of
DNA self-assembled networks scenario. For the same reason, we also
avoid considering solutions based on hardware-redundancy to
dynamically recover defects as in~\cite{constantinides2006, kohler2010, kim2006, park2006, ebrahimi2013}.

In~\cite{mejia_ipdps06} authors present SR, an approach the solves these
limitations by allowing turn restrictions to be placed locally,
independently from other restrictions. The whole network is
partitioned into segments, and each bidirectional turn
restriction can be freely chosen within a segment in order to guarantee
deadlock freedom and connected networks. This \emph{locality
independence} property, together with no requirement of any particular tree/root
choice, would make it the best choice for the given scenario;
however, its topology independency still requires the knowledge of the
whole network graph in order to find the segments.

The \disr{} approach presented in this work is a first attempt of
reaching the same goals of a segment based approach, that is
establishing segments, without requiring any topology graph or
centralized execution.

>>>>

\section{Related Works and Contribution}
\label{sec:related_works}
Several works~\cite{sancho2002, skeie2002, skeie2004, koibuchi2003} have been presented addressing the problem of topology
agnostic routing. While they show interesting
performances, the requirement for virtual channels resources 
to avoid deadlock makes them not suitable for the specific DNA self
assembled networks we are assuming as background.
Other topology independent algorithms~\cite{schroeder1991, koibuchi2001, cherkasova1996} do not use virtual channels to
achieve deadlock freedom, but turn prohibitions. 
In particular, authors of~\cite{Patwardhan05evaluatingthe} exploits the creation of a spanning tree of the
topology, then placing bidirectional restrictions by avoiding a packet
to traverse the same link in both up and down directions.
While the hierarchical nature of this approach can lead to uneven traffic
distribution, with many packets traversing upper links (near to the
root), this is quite acceptable in classical wide area networks
topologies with a limited number of nodes. Other approaches such as
FX~\cite{sancho2000} mitigate this
issue, but the set of turn restrictions is still prefixed,
strictly depending on the particular tree root selected. 
Other solutions try to approach the issue of irregular
topologies by limiting the number~\cite{dally1994, duato1997, gomez2004, koibuchi2008} or the
location~\cite{zhang2008, sui2000, flich2008, liu2011} of missing links. This
restriction is clearly unacceptable given the hight-defect rates of
DNA self-assembled networks scenario. For the same reason, we also
avoid considering solutions based on hardware-redundancy to
dynamically recover defects as in~\cite{constantinides2006, kohler2010, kim2006, park2006, ebrahimi2013}.

In~\cite{mejia_ipdps06} authors present an approach the solves these
limitations by allowing turn restrictions to be placed locally,
independently from other restrictions. The whole network is
partitioned into segments, and a bidirectional turn
restriction can be freely chosen within a segment in order to guarantee
deadlock freedom while preserving network connectivity. This \emph{locality
independence} property, avoiding the requirement of any particular tree/root
choice, would make it the best choice for the given scenario;
however, it still requires the knowledge of the
whole network graph in order to find the segments.

The \disr{} approach presented in this work is a first attempt of
reaching the same goals of a segment based turn prohibition  without
requiring any topology graph or centralized execution.  In particular
\disr{} aims to obtain a segment network partitioning using a
completely different mechanism, based on a distributed exchange of
small setup packets. The following main features distinguish \disr{}
from classic approaches to segment partitioning: 
\begin{itemize}
\item No centralized entity is globally responsible and/or aware of
what is going on, i.e. the status of the \disr{} execution is
collectively distributed among the nodes
\item  No defect map and/or topology graph is used as input,
thus, the topology has to be discovered \emph{while} segments are
created
\item At the end of the execution, no segment list is created or
stored anywhere: each node is only aware of belonging to some segment,
ignoring the presence of other nodes in the same segment and even the
presence of other segments in the network
\end{itemize}

### 2014-02-12 PAPER/simulation.tex
General text reorganization/fixes

% TODO: long paper
\subsection{Other Optimisations}
Some optimisation parameters, which demostrated to improve the \disr{}
results have been fixed to some good values, but are not subject to
further investigation; once again the focus here is not the
optimal set of segments, but just demostrating a working approach. 
These are optimisations are:
\begin{itemize}
\item{cycle\_links}: max number of retries across the set of links of
each node. While searching for a free link due an incoming request,
the same request is cancelled after a given number of tries. This
gives to preceding node change to test a different paths. Default is set to 1.
\item{bootstrap\_timeout}: number of time units that a bootstrap node
should wait before assuming that a livelock in the starting segment
process has occurred. In the worst case, we can imagine that longest
path required is the one returning to the bootstrap node after having
travelled across all the links. So, althought this is just an extreme
situation, a good upper limit can be safely be set to NxN.
\item{bootstrap\_immunity}: in order to avoid the failure of the whole \disr{}
setup process, a bootstrap node should not have defective links.
Enabling this optimisation, a bootstrap node is immune to defects.
We may think of a pre-bootstrap phase that properly selects (from upper
layer via) a bootstrap node which is tested as properly connected. We
enabled this optimisation, however empirical tests showed us that only
simulations using bootstrap nodes placed on edges would be heavily
affected, since these are nodes starting with a lower number of links,
e.g. corner nodes could only have two connected direction, so even a
defective link could prevent starting segment packet to return to
bootstrap to close the loop a create the segment.
\end{itemize}

@@

>>>>

@@

% TODO: long paper
\subsection{Other Optimisations}
Some optimisation parameters, which demostrated to improve the \disr{}
results have been fixed to some good values, but are not subject to
further investigation; once again the focus here is not the
optimal set of segments, but just demostrating a working approach. 
These are optimisations are:
\begin{itemize}
\item \emph{cycle\_links}: max number of retries across the set of links of
each node. While searching for a free link due an incoming request,
the same request is cancelled after a given number of tries. This
gives to preceding node change to test a different paths. Default is set to 1.
\item \emph{bootstrap\_timeout}: number of time units that a bootstrap node
should wait before assuming that a livelock in the starting segment
process has occurred. In the worst case, we can imagine that longest
path required is the one returning to the bootstrap node after having
travelled across all the links. So, althought this is just an extreme
situation, a good upper limit can be safely be set to NxN.
\item \emph{bootstrap\_immunity}: in order to avoid the failure of the whole \disr{}
setup process, a bootstrap node should not have defective links.
Enabling this optimisation, a bootstrap node is immune to defects.
We may think of a pre-bootstrap phase that properly selects (from upper
layer via) a bootstrap node which is tested as properly connected. We
enabled this optimisation, however empirical tests showed us that only
simulations using bootstrap nodes placed on edges would be heavily
affected, since these are nodes starting with a lower number of links,
e.g. corner nodes could only have two connected direction, so even a
defective link could prevent starting segment packet to return to
bootstrap to close the loop a create the segment.
\end{itemize}
